---
title: "p8105_hw2_yz4433"
author: "Yifei Zhao"
date: "2022-10-05"
output: github_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
library(tidyverse)
library(readxl)
```

## Problem 1
### read and clean data
```{r}
trans_ent = read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

### Distinct Station
```{r, results = 'hide'}
trans_ent %>% 
  select(station_name, line) %>% 
  distinct
```

### ADA compliant
```{r, results = 'hide'}
trans_ent %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

### Proportion of station entrances or exits without vending allow entrance
```{r}
trans_ent %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

### Reformat data and do calculations
```{r, results = 'hide'}
trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct
trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

## Problem 2
### Manipulate the Mr.Trash Wheel sheet
```{r}
data_mr = read_excel("data/Trash Wheel Collection Data.xlsx") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(group = "mr")
data_mr = data_mr[,-15:-16]
data_mr[,13] = as.integer(unlist(round(data_mr[,13])))
```

### Manipulate the Professor Trash Wheel sheet and make combination
```{r}
data_pro = read_excel("data/Trash Wheel Collection Data.xlsx",2) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  add_column(sports_balls = 0, .after = "chip_bags") %>% 
  mutate(group = "professor")
data_pro[,3] = as.character(unlist(data_pro[,3]))
# Combinationa and get data_trash
data_trash =
  bind_rows(data_mr, data_pro) %>%
  janitor::clean_names() %>%
  select(group, everything())
```
For Mr. Trash Wheel data, a `r nrow(data_mr)` $\times$ `r ncol(data_mr)` dataset, the key variables are date and weight_tons, and it illustrates the amount of total litter and litter type from May 2014 to July 2022. For Professor Trash Wheel data, a `r nrow(data_pro)` $\times$ `r ncol(data_pro)` dataset, the key variables are also date and weight_tons, and it illustrates the amount of  precipitation and litter type from Jan 2017 to July 2022. The final resulting dataset has 641 observations. The key variables are date and weight_tons. After calculation, the total weight of trash collected by Professor Trash Wheel is `r sum(data_pro$weight_tons)`, and the total number of sports balls collected by Mr. Trash Wheel in 2020 is `r sum(data_mr$sports_balls[381:452])`.

## Problem 3
### pols-month manipulation
```{r}
data_pols = 
  read_csv("data/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day")) %>%
  mutate(president = "group")
data_pols$president = ifelse(data_pols$prez_gop != 0, 'gop', 'dem') 
data_pols = select(data_pols, -day, -prez_dem, -prez_gop)
data_pols$month = month.abb[as.numeric(unlist(data_pols$month))]
```

### snp manipulation
```{r}
data_snp = 
  read_csv("data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
separate(date, into = c("month", "day", "year")) %>%
select(year, month, close, -day)
data_snp$month = month.abb[as.numeric(unlist(data_snp$month))]
data_snp$tw = c('20')
data_snp$ni = c('19')
data_snp[1:187,] = mutate(data_snp[1:187,], year = str_c(tw,year))
data_snp[188:787,] = mutate(data_snp[188:787,], year = str_c(ni,year))
data_snp = data_snp[,-4:-5]
```

### unemployment manipulation
```{r}
data_unem = 
  read_csv("data/fivethirtyeight_datasets/unemployment.csv") %>%
  rename(year = Year) %>% 
  pivot_longer(
    Jan:Dec,
    names_to = "month", 
    values_to = "value")
data_unem[,1] = as.character(unlist(data_unem[,1]))
data_unem = data_unem[-811:-816,]
```

### Merge datasets
```{r}
middata = full_join(data_pols, data_snp, by = c("year","month"))
finaldata = full_join(middata, data_unem, by = c("year","month"))
```
For pols-month data, a `r nrow(data_pols)` $\times$ `r ncol(data_pols)` dataset, the key variables are year, month, and president, and it illustrates the statistics of national politicians from Jan 1947 to Jun 2015. During this period, the president is republican for 390 months, and the president is democratic for 432 months. For snp data, a `r nrow(data_snp)` $\times$ `r ncol(data_snp)` dataset, the key variables are year, month, and close, and it illustrates the statistics of the closing value of Standard & Poor’s stock market on the associate date from Jan 1950 to Jul 2015. In total, the closing values are in the descending trend  totally. For unemployment data, a `r nrow(data_unem)` $\times$ `r ncol(data_unem)` dataset, the key variables are year, month, and value, and it illustrates the percentage of unemployment monthly from Jan 1948 to Jun 2015. The maximum values are `r max(data_unem$value)` in Nov and Dec 1982. Also we find the maximum average percentage of unemployment by year is about 9.7 in 1982. In contrast, the minimum values are `r min(data_unem$value)` in May and Jun in 1953. Also we find the minimum average percentage of unemployment by year is about 2.9 in 1953. After combination these three datasets, from the resulting dataset, a `r nrow(finaldata)` $\times$ `r ncol(finaldata)` dataset, the key variables are year, month, president, close, and value, and it illustrates the figures of politics, stock market, and employment by month or by year to some aspects from Jan 1947 to Jul 2015. We apply funjoin to guarantee that the finaldata involves features of each dataset discussed above. 